COVID-19 and Online Learning Tools

### 2021-12-28 Paper_online_Learning.tex
Updated

Even though distance education has a long history that goes back almost two centuries with significant modifications,
alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}. Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.

@@

According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and the need for urgent responses.

@@

\hypertarget{background}{%
\section{Background}\label{background}}

@@

According to \textcite{moore2011learning}, ``distance education'' is the most renowned term used when referencing distance learning. However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version,'' ``newer version'' and ``improved version'' of distance learning. The Google Trend search patterns correspond to the term ``online learning'' confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)). Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``distance learning'' and ``online teaching'' demonstrate a similar trend during the study period. COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases. A similar decline can be observed in distance education-related terms (Figures \ref{fig:distanceLearningWorldAnalysis}). This confirms the life cycle of Google Trend search patterns introduced in \textcite{jarynowski2020perception}. Further, a second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic. This could be due to the nearly total closures of schools that happened with the declaration of COVID-19 as a pandemic.

@@

The degree of digression of the blue line from the ideal 45-degree straight line indicates a dissimilarity between the two time
series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)). In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was known as ``corona'', it was not a global concern and was mostly limited to China. During that time, people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic. This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).

@@

Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable. These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic. Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has attracted considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)). `Udemy' has got the second highest attention. However, `Canvas' has received far more attention than `Udemy'. Under self-directed learning content, `Quizlet' has received considerably high attention throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)). It provides learning flashcards, an AI based learning assistant, and games to support student learning in multiple subjects and languages. The second highest option only provides support for language learning. Quizlet's ability to provide support for multiple subjects might be the reason for its ability to outperform its competitors in the market. Under mobile reading applications, `Reads', which provides digital stories with illustrations in multiple languages has consistently been the most popular option (Figure \ref{fig:plot2} (f)). Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had received a lot of attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)). However, `WhatsApp' has outperformed `Zoom' in attracting public attention during the pandemic. In contrast, global attention towards `Zoom' has decreased over time. `Teams' has also gotten considerable attention in comparison to the other remaining options available in the market. During the pandemic, tools for teachers to create digital learning content, such as ``Trello,'' which provides a visual collaboration platform for teachers for coursework planning, classroom organization, and collaboration, received a lot of attention. `Nearpod', which promotes active learning and student engagement through informative and interactive learning activities has also gotten considerable attention during the pandemic (Figure \ref{fig:plot2} (h)). A significant peak can be observed in the `Nearpod' series, during the first academic break. Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got significant global attention (Figure \ref{fig:plot2} (i)). `Brookings' provides a large catalogue of learning innovations, though its focus is not limited to distance learning solutions. `UNHCR' also provides an extensive list of distance learning solutions from the United Nations agency for refugees. Student assessment is vital in ensuring the quality of education. However, due to an unexpected and sudden school closure, administering examinations at a distance became a major concern. Under tools for online proctoring, `Pearson VUE' which provides computer based testing services in

>>>>

Even though distance education has a long history that goes back almost two centuries with significant modifications,
alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}. Sudden unexpected movement from classrooms to home schooling at large-scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.

@@

According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions \autocite{unesco2020DLsolutions} that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and the need for urgent responses.

@@

\hypertarget{related-work}{%
\section{Related work}\label{related-work}}

@@

According to \textcite{moore2011learning}, ``distance education'' is the most renowned term used when referencing distance learning. However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version,'' ``newer version'' and ``improved version'' of distance learning. The Google Trend search patterns correspond to the term ``online learning'' confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)). Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``distance learning'' and ``online teaching'' demonstrate a similar trend during the study period. COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases. A similar decline can be observed in distance education-related terms (Figures \ref{fig:distanceLearningWorldAnalysis}). This confirms the life cycle of Google Trend search patterns introduced in \textcite{jarynowski2020perception}. Further, the second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic. This could be due to the nearly total closures of schools that happened with the declaration of COVID-19 as a pandemic.

@@

The degree of digression of the blue line from the ideal 45-degree straight line indicates a dissimilarity between the two time
series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)). In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide \autocite{ciotti2019covid}. When the virus was known as ``corona'', it was not a global concern and was mostly limited to China. During that time, people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic. This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).

@@

Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as a tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable. These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic. Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has attracted considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)). `Udemy' has got the second highest attention. However, `Canvas' has received far more attention than `Udemy'. Under self-directed learning content, `Quizlet' has received considerably high attention throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)). It provides learning flashcards, an AI based learning assistant, and games to support student learning in multiple subjects and languages. The second highest option only provides support for language learning. Quizlet's ability to provide support for multiple subjects might be the reason for its ability to outperform its competitors in the market. Under mobile reading applications, `Reads', which provides digital stories with illustrations in multiple languages has consistently been the most popular option (Figure \ref{fig:plot2} (f)). Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had received a lot of attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)). However, `WhatsApp' has outperformed `Zoom' in attracting public attention during the pandemic. In contrast, global attention towards `Zoom' has decreased over time. `Teams' has also gotten considerable attention in comparison to the other remaining options available in the market. During the pandemic, tools for teachers to create digital learning content, such as ``Trello,'' which provides a visual collaboration platform for teachers for coursework planning, classroom organization, and collaboration, received a lot of attention. `Nearpod', which promotes active learning and student engagement through informative and interactive learning activities has also gotten considerable attention during the pandemic (Figure \ref{fig:plot2} (h)). A significant peak can be observed in the `Nearpod' series, during the first academic break. Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got significant global attention (Figure \ref{fig:plot2} (i)). `Brookings' provides a large catalogue of learning innovations, though its focus is not limited to distance learning solutions. `UNHCR' also provides an extensive list of distance learning solutions from the United Nations agency for refugees. Student assessment is vital in ensuring the quality of education. However, due to an unexpected and sudden school closure, administering examinations at a distance became a major concern. Under tools for online proctoring, `Pearson VUE' which provides computer based testing services in

### 2021-12-28 main.Rmd
Updated

Even though distance education has a long history that goes back  almost two centuries with significant  modifications,
alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].  Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher]. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.

@@

According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and the need for urgent responses.

@@

# Background

@@

According to @moore2011learning, "distance education" is the most renowned term used when referencing distance learning. However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version," "newer version" and "improved version" of distance learning. The Google Trend search patterns correspond to the term "online learning" confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)). Despite quantitative differences in terms of relative search volume, the series generated by the two terms "distance learning" and "online teaching" demonstrate a similar trend during the study period. COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases. A similar decline can be observed in distance education-related terms  (Figures \ref{fig:distanceLearningWorldAnalysis}). This confirms the life cycle of Google Trend search patterns introduced in @jarynowski2020perception. Further,  a second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic. This could be due to the nearly total closures of schools that happened  with the declaration of  COVID-19 as a pandemic.

@@

The degree of digression of the blue line from the ideal 45-degree straight line indicates a dissimilarity between the two time
series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)). In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide [@ciotti2019covid].When the virus was known as "corona", it was not a global concern and was mostly limited to China. During that time, people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic. This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).

@@

Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable. These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic. Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has attracted considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)). 'Udemy' has got the second highest attention. However, 'Canvas' has received far more attention than 'Udemy'. Under self-directed learning content, 'Quizlet' has received considerably high attention  throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)). It provides learning flashcards, an AI based learning assistant, and games to support student learning in multiple subjects and languages. The second highest option only provides  support for language learning.  Quizlet's ability to provide support for multiple subjects might be the reason for its ability to outperform its  competitors in the market. Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently been the most popular option (Figure \ref{fig:plot2} (f)). Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had received a lot of attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)). However, 'WhatsApp' has outperformed 'Zoom'  in attracting public attention during the pandemic. In contrast, global attention towards 'Zoom'  has decreased over time. 'Teams' has also gotten considerable attention in comparison to the other remaining options available in the market. During the pandemic, tools for teachers to create digital learning content, such as "Trello," which provides a visual collaboration platform for teachers for coursework planning, classroom organization, and collaboration, received a lot of attention. 'Nearpod', which promotes active learning and student engagement through informative and interactive learning activities has also gotten considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).  A significant peak can be observed in the 'Nearpod' series, during the first academic break. Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got significant global attention (Figure \ref{fig:plot2} (i)). 'Brookings' provides a large catalogue of learning innovations, though its focus is not limited to distance learning solutions.  'UNHCR' also  provides an extensive list of distance learning solutions from the United Nations agency for refugees. Student assessment is vital in ensuring the quality of education. However, due to an unexpected and sudden school closure, administering examinations at a distance became a major concern. Under tools for online proctoring, 'Pearson VUE' which provides computer based testing services in

>>>>

Even though distance education has a long history that goes back  almost two centuries with significant  modifications,
alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].  Sudden unexpected movement from classrooms to home schooling at large-scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher]. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.

@@

According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions [@unesco2020DLsolutions] that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and the need for urgent responses.

@@

# Related work

@@

According to @moore2011learning, "distance education" is the most renowned term used when referencing distance learning. However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version," "newer version" and "improved version" of distance learning. The Google Trend search patterns correspond to the term "online learning" confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)). Despite quantitative differences in terms of relative search volume, the series generated by the two terms "distance learning" and "online teaching" demonstrate a similar trend during the study period. COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases. A similar decline can be observed in distance education-related terms  (Figures \ref{fig:distanceLearningWorldAnalysis}). This confirms the life cycle of Google Trend search patterns introduced in @jarynowski2020perception. Further,  the second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic. This could be due to the nearly total closures of schools that happened  with the declaration of  COVID-19 as a pandemic.

@@

The degree of digression of the blue line from the ideal 45-degree straight line indicates a dissimilarity between the two time
series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)). In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide [@ciotti2019covid]. When the virus was known as "corona", it was not a global concern and was mostly limited to China. During that time, people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic. This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).

@@

Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  a tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable. These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic. Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has attracted considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)). 'Udemy' has got the second highest attention. However, 'Canvas' has received far more attention than 'Udemy'. Under self-directed learning content, 'Quizlet' has received considerably high attention  throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)). It provides learning flashcards, an AI based learning assistant, and games to support student learning in multiple subjects and languages. The second highest option only provides  support for language learning.  Quizlet's ability to provide support for multiple subjects might be the reason for its ability to outperform its  competitors in the market. Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently been the most popular option (Figure \ref{fig:plot2} (f)). Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had received a lot of attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)). However, 'WhatsApp' has outperformed 'Zoom'  in attracting public attention during the pandemic. In contrast, global attention towards 'Zoom'  has decreased over time. 'Teams' has also gotten considerable attention in comparison to the other remaining options available in the market. During the pandemic, tools for teachers to create digital learning content, such as "Trello," which provides a visual collaboration platform for teachers for coursework planning, classroom organization, and collaboration, received a lot of attention. 'Nearpod', which promotes active learning and student engagement through informative and interactive learning activities has also gotten considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).  A significant peak can be observed in the 'Nearpod' series, during the first academic break. Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got significant global attention (Figure \ref{fig:plot2} (i)). 'Brookings' provides a large catalogue of learning innovations, though its focus is not limited to distance learning solutions.  'UNHCR' also  provides an extensive list of distance learning solutions from the United Nations agency for refugees. Student assessment is vital in ensuring the quality of education. However, due to an unexpected and sudden school closure, administering examinations at a distance became a major concern. Under tools for online proctoring, 'Pearson VUE' which provides computer based testing services in
